Data Specialization for a Staged Lambda Calculus

### 2014-11-06 paper/splitting.tex
Editing cut at 4.0 and the intro paragraph of 4.1.

Although terms in \lang\ clearly distinguish their stage \bbone\ and stage
\bbtwo\ components, these components may be interleaved in complex ways. The
goal of \emph{stage splitting} is to produce from a term in \lang\ a pair of
terms in an unstaged language \langmono, which respectively perform all the
stage \bbone\ and stage \bbtwo\ subcomputations of the original term.

Since splitting stage \bbtwo\ terms is easier than splitting stage \bbone\
terms, we begin by describing how to split stage \bbtwo\ terms.

@@

Splitting a stage \bbtwo\ term $\coltwo e A$ produces a \emph{precomputation}
term $p$ and a \emph{resumer} function $l.r$, which contain all stage \bbone\
and \bbtwo\ subcomputations of $e$, respectively. As stage \bbtwo\
computations within $e$ can depend on its stage \bbone\ computations, 
$r$ must depend on the results of $p$, which splitting accomplishes via the variable $l$. 
Given this form, the goal for splitting is to ensure that evaluating
$\letin l p r$ agrees with the value to which $e$ itself evaluates.

\ur{Correctness will have
  to go beyond this, by saying for example that all stage 1 parts are
  in p etc.  Note that this correctness allows a trivial split
   that shoves everything to r, e.g., p = () and r is ``e''
  (might need to flatten).}

>>>>

Although terms in \lang\ clearly distinguish their stage \bbone\ and stage
\bbtwo\ components, these components may be interleaved in \lang\ code. The
goal of \emph{stage splitting} is to statically separate the stage \bbone\ and \bbtwo\ subcomputations
of a term $e$ in \lang\ into a pair of
terms $e_1$ and $e_2$ in an unstaged language \langmono, where $e_1$ performs all the
stage \bbone\ subcomputations in $e$ and $e_2$ performs all of the stage \bbtwo\ subcomputations.
Intuitively, splitting must ensure that running $e_2$ on the output of $e_1$ yields the same value 
as evaluating $e$ in \lang.

Since the mechanics of splitting stage \bbtwo\ terms are conceptually simpler than those for stage \bbone\
terms, we begin by describing how to split stage \bbtwo\ terms before proceeding to more complicated cases.

@@

Splitting a stage \bbtwo\ term $\coltwo e A$ generates
 a \emph{precomputation} $p$, and a \emph{resumer} function $l.r$. Whereas direct evaluation of $e$ in \lang\ reduces the stage \bbone\ subcomputations
of $e$ in place to create a residual, splitting statically isolates these computations into
$p$.  The resumer accepts the value produced by the precomputation as input
% (bound to the free variable $l$)
and executes all stage 2 subcomputations.   
Specifically, if $e$ in \lang\ reduces to the final value $v$, 
then after splitting the term $\letin l p r$ should reduce to the same value in \langmono.
(The resumer, bound to the output of $p$, is equivalent to the residual term defined in~\ref{ssec:dynamics})

%	\ur{Correctness will have
%	  to go beyond this, by saying for example that all stage 1 parts are
%	  in p etc.  Note that this correctness allows a trivial split
%	   that shoves everything to r, e.g., p = () and r is ``e''
%	  (might need to flatten).}
